---
title: "a3_task3_sam_fearer"
author: "Sam Fearer"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE, echo = TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(scales)
```

## SECTION ONE
## Read in the urchin data and clean names

```{r}
urchin = read_csv(here("data","urchins.csv")) %>% 
  clean_names()
  
```


## Exploratory data viz for comparison of red and purple urchins, by size

```{r}

#histogram
ggplot(data = urchin, (aes(x = size))) +
  geom_histogram() + 
  facet_wrap(~common_name)

#boxplot
ggplot(data = urchin, (aes(x = common_name, y = size))) +
  geom_boxplot()
```


## Descriptive statistics table (presenting: mean, median, standard deviation, sample size)

```{r}

urchin_stats <- urchin %>% 
  group_by(common_name) %>% 
  summarize(mean_size = mean(size),
            median_size = median(size),
            sd_size = sd(size),
            sample_size = n())

```

## Size comparison, using statistics from graphs and from "urchin_stats":
> The red urchin sample had a larger mean and median size (roughly 2.3cm, or 40%, larger) than that of purple urchins. However, the red urchin sample had a smaller sample size (approximately 15% smaller), and more viariance from the mean (the standard deviation of the red urchin set was almost 40% larger than that of the purple urchin set). Given these figures, we can understand the red urchins in question to have a larger average size, but more variability between individual samples. That said, it is the red urchin dataset that most resembles a normal distribution. 

## SECTION TWO

## Explore purple urchin data by site and year

```{r}

purple_urchins <- urchin %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year_new = year(date_new)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"
  ) )


```

## Summary table and graph for purple urchin data

```{r}
#summary table
#purple_urchin_summary <- purple_urchins %>% 
#  select("year_new", "site_full")

  
#graph from table data
#ggplot(data = purple_urchin_summary, aes(x = year_new, color = site_full)) +
 # geom_histogram(bins = 30) +
 # facet_wrap(~site_full)



#summary table
purple_urchin_summary <- purple_urchins %>% 
  select("year_new", "site_full") %>% 
  count(site_full, year_new)

  
#graph from table data
ggplot(data = purple_urchin_summary, aes(x = year_new, y = n, color = site_full)) +
  geom_col(bins = 30) +
  facet_wrap(~site_full)
  
  
```


























